Estimating the magnetization distribution within rectangular rock samples

### 2016-07-22 manuscript/manuscript.tex

Figure \ref{fig:estimate-sensor2sample} shows the four
estimated magnetization distributions produced in this
test. All these estimates were obtained by using a regularization
parameter $\mu_{0} = 1 \times 10^{-10}$ (equation \ref{eq:goal-function}).
This figure also shows that,
as in the previous tests, the estimate obtained by 
inverting the full data set (black dots and black dashed lines) 
best approximates the true magnetization distribution
(red dots and red lines). However, even this estimate
did not completely retrieved the magnetization distribution of the synthetic
sample (red dots).
Notice that the observed data (Figures \ref{fig:datafit-sensor2sample}a, 
d, g and j) present striations aligned with the acquisition lines
on each observation plane. These striations are produced by the
presence of position noise in the observation points.
Despite these striations, the (almost good) estimated magnetization 
distribution obtained by inverting the full data set 
(black dots and black dashed lines in Figure \ref{fig:estimate-sensor2sample}) 
yields predicted data (Figures \ref{fig:datafit-sensor2sample}b, e, h and k) 
that recover the observed data (Figures \ref{fig:datafit-sensor2sample}a, 
d, g and j) in an acceptable manner.
This acceptable data fit is shown by the normalized histograms of 
the residuals between the predicted and
observed data (Figures \ref{fig:datafit-sensor2sample}c, f, i and l).
The sample means $\mu$ are close to zero but the sample standard 
deviations $\sigma$ present some discrepancies with respect to
that one of the pseudo-random noise in magnetic data.
According to these histograms, the poorest data fit occurs in the 
observation plane $\alpha = 3$ (Figure \ref{fig:datafit-sensor2sample}l),
where the sample standard deviation $\sigma \approx 60$~$\mu$T.

@@

The method proposed by \citet{oda2011} 
uses SQUID microscopy to identify the boundaries of fine magnetic stripes 
(smaller than $1$~mm) with approximately reversed magnetization. These magnetic stripes are 
then correlated with a standard magneto-stratigraphic time scale, providing a tool for 
estimating ages and growth rates for hydro-genetic ferro-manganese crusts with unprecedented 
spatial resolution.
\citet{oda2011} applied their method to analyse a block of ferro-manganese crust obtained 
from a seamount in the Northwest Pacific Ocean. They have cut and sliced a columnar block 
parallel to the growth lamination at $1.5$~mm intervals using a $0.3$-mm-thick diamond-wire 
saw. After that, the magnetization of the slices have been estimated by using a 
SQUID magnetometer.
Note that, to determine the magnetization along the sample, it was necessary to slice 
it in small parts. It would be useful instead to determine the bulk magnetization along 
the sample by using a non-destructive method.
Here, we present a feasibility study aiming at  
determining the magnetization distribution along a simulated sample
by applying our method and directly inverting the high-resolution measurements of 
the magnetic induction around the sample.

@@

The first one concerns the estimated magnetization intensities 
(Figure \ref{fig:estimate-real}a).
As we can see, the estimated intensities exhibit an interesting pattern 
formed by a well-defined cycle of a low value that is followed by three higher 
values. This cycle repeats four times,
coinciding with the number of prisms forming our synthetic
sample.
For convenience, grouped the estimated values representing
the magnetization within each prism.
The low values observed in the estimated magnetization intensities
(Figure \ref{fig:estimate-real}a), as well
as the corresponding inclination and declination values
(Figures \ref{fig:estimate-real}b and c), are represented by black
triangles.
This remarkable pattern is consistent with the 
magnetite precipitation during the manufacturing of our sample.
This precipitation would decrease the magnetite concentration at the
top of each sample prism, increasing the concentration towards
their bottom.
The magnetic induction produced by these low-concentration zones
there must be superimposed by those ones produced by the 
high-concentration zones, which makes the correct estimation 
of the magnetization direction in these low-concentration zones 
very difficult.
Due to this possible precipitation effect, 
we did not take into consideration the estimated values that coincides
with the low-concentration zones 
(black triangles in Figure \ref{fig:estimate-real}). Consequently,
we based our interpretation on the estimated values that are 
not located at the low-concentration zones.

@@

These results show that our method could be an interesting complement
of traditional paleomagnetic techniques aiming at providing a 
non-destructive diagnostic of geological samples.
Nevertheless, further tests using either manufactured and geological samples 
need to be carried out with the purpose of validating our approach,
better evaluating the effects of position noise
and misalignment problems during the scanning stage and
also developing automatic pre-processing techniques.

@@

 \begin{figure}
 \noindent \includegraphics[width=20pc]{Figs/Fig13_HQ.pdf}
 \caption{Manufactured sample. 
 (a) Four prisms forming the sample. The IRM magnetization of these
 prisms are approximately parallel to the vertical planes represented
 in grey.
 (b) Magnetization (thick arrows) of the prisms on the vertical 
 planes shown in (a).
 (c) Resultant sample obtained by juxtaposing the magnetized prisms.
 The numbers indicate the index of each prism, whose magnetization
 is represented by the thick arrows. The inclination and declination
 values of the IRM within each prism is shown in Table 
 \ref{tab:IRM-real-sample}.
 The resulting sample is referred to a MCS (Figure \ref{fig:sample}) 
 with origin represented by the black dot and axes $x$ and $z$ 
 represented by the thin arrows.}
 \label{fig:real-sample}
 \end{figure}

>>>>

Figure \ref{fig:estimate-sensor2sample} shows the four
estimated magnetization distributions produced in this
test. All these estimates were obtained by using a regularization
parameter $\mu_{0} = 1 \times 10^{-10}$ (equation \ref{eq:goal-function}).
This figure also shows that,
as in the previous tests, the estimate obtained by 
inverting the full data set (black dots and black dashed lines) 
best approximates the true magnetization distribution
(red dots and red lines). However, even this estimate
did not completely retrieved the magnetization distribution of the synthetic
sample (red dots).
Notice that the observed data (Figures \ref{fig:datafit-sensor2sample}a, 
d, g and j) present striations aligned with the acquisition lines
on each observation plane. These striations are produced by the
presence of position noise in the observation points.
Despite these striations, the (almost good) estimated magnetization 
distribution obtained by inverting the full data set 
(black dots and black dashed lines in Figure \ref{fig:estimate-sensor2sample}) 
yields predicted data (Figures \ref{fig:datafit-sensor2sample}b, e, h and k) 
that recover the observed data (Figures \ref{fig:datafit-sensor2sample}a, 
d, g and j) in an acceptable manner.
This acceptable data fit is shown by the normalized histograms of 
the residuals between the predicted and
observed data (Figures \ref{fig:datafit-sensor2sample}c, f, i and l).
The sample means $\mu$ are close to zero but the sample standard 
deviations $\sigma$ present some discrepancies with respect to
that one of the pseudo-random noise in magnetic data.
According to these histograms, the poorest data fit occurs in the 
observation plane $\alpha = 3$ (Figure \ref{fig:datafit-sensor2sample}l),
where the sample standard deviation $\sigma \approx 47$~$\mu$T.

@@

\citet{oda2011} proposed a new method for determining absolute 
ages and growth rates for ferro-manganese crusts.
They applied their method to analyse a ferro-manganese crust sample 
obtained from a seamount in the Northwest Pacific Ocean and
showed that magnetostratigraphy using state of the art SQUID
microscopy is a promising chronological tool.
The method proposed by \citet{oda2011} uses SQUID microscopy 
on thin sections to identify the boundaries of fine magnetic stripes 
(smaller than $1$~mm) with approximately reversed magnetization. 
These magnetic stripes are then correlated with a standard 
magneto-stratigraphic time scale, providing a tool for 
estimating ages and growth rates for hydro-genetic 
ferro-manganese crusts with unprecedented 
spatial resolution.

\citet{oda2011} also used SQUID magnetometer measurements
to analyse a columnar block of their ferro-manganese crust sample.
They have cut and sliced the columnar block parallel to the growth 
lamination at $1.5$~mm intervals using a $0.3$-mm-thick diamond-wire 
saw. After that, they estimated the magnetization of the slices by 
using a SQUID magnetometer.
Note that, to determine the magnetization along the sample, 
it was necessary to slice it in small parts. It would be useful instead 
to determine the bulk magnetization along 
the sample by using a non-destructive method.
Here, we present a feasibility study aiming at  
determining the magnetization distribution along a simulated sample.
This feasibility study consists in applying our method and 
directly inverting simulated high-resolution measurements of 
the magnetic induction around the sample.

@@

The first one concerns the estimated magnetization intensities 
(Figure \ref{fig:estimate-real}a).
As we can see, the estimated intensities exhibit an interesting pattern 
formed by a well-defined cycle of a low value that is followed by three higher 
values. This cycle repeats four times,
coinciding with the number of prisms forming our synthetic
sample.
For convenience, grouped the estimated values representing
the magnetization within each prism.
The low values observed in the estimated magnetization intensities
(Figure \ref{fig:estimate-real}a), as well
as the corresponding inclination and declination values
(Figures \ref{fig:estimate-real}b and c), are represented by black
triangles.
This remarkable pattern is consistent with the 
magnetite precipitation during the manufacturing of our sample.
This precipitation would decrease the magnetite concentration at the
top of each sample prism, increasing the concentration towards
their bottom.
Figures \ref{fig:real-sample}b and c represent the magnetite 
precipitation as a gradient fill. The resultant
high-concentration zones at the bottom of the prisms forming the 
sample are represented in dark grey while the resultant
low-concentration zones at the top of the prisms are represented
in white. 
The magnetic induction produced by these low-concentration zones
there must be superimposed by those ones produced by the 
high-concentration zones, which makes the correct estimation 
of the magnetization direction in these low-concentration zones 
very difficult.
Due to this possible precipitation effect, 
we did not take into consideration the estimated values that coincides
with the low-concentration zones 
(black triangles in Figure \ref{fig:estimate-real}). Consequently,
we based our interpretation on the estimated values that are 
not located at the low-concentration zones.

@@

These results show that our method could be an interesting complement
to traditional paleomagnetic techniques aiming at providing a 
non-destructive diagnostic of geological samples.
Nevertheless, further tests using either manufactured and geological samples 
need to be carried out with the purpose of validating our approach,
better evaluating the effects of position noise
and misalignment problems during the scanning stage and
also developing automatic pre-processing techniques.
A further study need to be carried out in order to evaluate
the influence of, for example, the scanning geometry and the 
sensitivity of the magnetic sensor on the performance of our method.

@@

 \begin{figure}
 \noindent \includegraphics[width=20pc]{Figs/Fig13_HQ.pdf}
 \caption{Manufactured sample. 
 (a) Four prisms forming the sample. The IRM magnetization of these
 prisms are approximately parallel to the vertical planes represented
 in grey.
 (b) Magnetization (thick arrows) of the prisms on the vertical 
 planes shown in (a). The vertical gradient fill (shown in grey)
 represents the magnetite precipitation during the hardening of
 the sample. The resultant low-concentration zones at the top of
 the prisms are represented in white and the high-concentration zones
 at the bottom of the prisms are represented in dark grey.
 (c) Resultant sample obtained by juxtaposing the magnetized prisms.
 The numbers indicate the index of each prism, whose magnetization
 is represented by the thick arrows. The inclination and declination
 values of the IRM within each prism is shown in Table 
 \ref{tab:IRM-real-sample}.
 The resulting sample is referred to a MCS (Figure \ref{fig:sample}) 
 with origin represented by the black dot and axes $x$ and $z$ 
 represented by the thin arrows.}
 \label{fig:real-sample}
 \end{figure}
